---
title: "Computing Integrals Monte Carlo Way"
author: "Rahul Goswami"
date: "18/01/2022"
output: pdf_document
keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Rejection Sampling Method

Its always not easy to withdraw samples from posterior $\pi(\theta|x)$ distribution.Most of the time 
are not familiar with the functional form of the posterior distribution. 

Supoose we wanna take a sample from the posterior distribution $\pi(\theta|x)$
Then we will find another probability distribution $p(\theta)$ which have the following properties

1. Easy to withdraw samples from
2. Resembles the posterior distribution
3. For all parameter $\theta$ and a constant $k$ , $\pi(\theta|x) \leq k  p(\theta)$

### Algorithm

1. Take a sample from the from the distribution $p(\theta)$ and a Uniform Random Variable $U$
2. If $U < \frac{\pi(\theta|x)}{k \cdot p(\theta)}$ then accept the sample
3. If $U > \frac{\pi(\theta|x)}{k \cdot p(\theta)}$ then reject the sample


The Performance of the Rejection Sampling Method is measured by Acceptance Rate.

### Example

Suppose we want to withdraw samples from normal distribution with mean $\mu$ and variance $\sigma$,which equivalent to get samples from standard Normal distribution, because we just have to do a simple linear transformation to get a distribution with mean $\mu$ and variance $\sigma$.So we will be using the standard Normal distribution

Now we are taking proposaldensity or in some literature mentioned as candidate density $p(\theta)$ as an exponential distribution with mean $1$, while we know that exponential random variable is always positive and hence we will be taking the absolute value of the Standard Normal random variable, and then multiply iy by -1 by generating a uniform random variable $U$.Whenever $U$ is less than 0.5


$$
p(\theta) = e^{-\theta}  \\
\pi(\theta|x) = \frac{2}{2\pi}  e^{-\frac{1}{2}(\theta)^2}1_{x \geq0}
$$

Then


$\frac{\pi(\theta|x)}{p(\theta)}$ is the ratio of the posterior distribution and the candidate density.It will be at 
maximum at $\theta = 1$ thus k = $\sqrt{2e / \pi} \approx 1.32$ 

Then steps for generating samples from the posterior distribution are as follows
1. Take a sample from exponential distribution with mean $1$ and a uniform random variable U
2. If $U \leq \frac{\frac{1}{\sqrt{2\pi}}e^{-\frac{\theta^2}{2}}}{\sqrt{2e/\pi} e^{-\theta}} \ i.e \ U \leq e^{-(1 -\theta)^2/2}$ then accept the sample
3. Generate another uniform random variable $U$, if U is less than 0.5 then multiply the sample by -1



```{r RejectionSampling}
nsample = 10000
sample = c()
count = 0
while(length(sample) < nsample){
  U = runif(1)
  count = count + 1
  theta = rexp(1)
  U2 = runif(1)
  if(U <= exp((-(1-theta)^2)/2)){ 
    if(U2 <= 0.5){
      theta = -theta
    }
    
    sample = c(sample, theta)
  }
}
cat("Acceptance Rate: ", count/nsample)
plot(density(sample))
```


